
# Drug Sale Campaign
## About Drug Sale Campaign

When a new drug hits the market, not every clinician is aware of the drug treatment because there are too much information on the web these days. One way to promote a new drug awareness is through campaigns. This could be advertisements on certain medical related websites or targeted messages in certain medical software applications like a prescribing application. 

This project focuses on evaluating a campaign effectiveness to promote a new drug called 'FEELA-GOOD' . FEELA-GOOD is a medication that relieves pain and makes you feel good. This is a special drug because it has no side effects except maybe emptying your cash wallet.We are all aware of the danger of opioids in pain management. Opioids are great for relieving pain but there are high risk of addiction and overdose that might lead to death. According to CDC, in 2021, over 75% of the approximately 107,000 drug overdose fatalities were attributed to opioids.This translates to a staggering 80,000+ lives lost to opioid overdoses in the same year, equating to an average of 220 individuals per day. It's crucial to recognize that these figures are likely conservative estimates, and the opioid crisis is likely to escalate, as evidenced by numerous tragic overdose incidents reported in the news.


## About the Data set

This  synthetic dataset gives information about the doctors'  prescription behavior and whether the doctor has seen the promotional message or not. There are only 6 columns. We will use the data to analyze if the messaging had any positive lift in 'FEELA-GOOD' prescriptions. There will be visuals to view data in different segments.Finally a T-test will be used to see if the campaign effect is statistically significant.


* Attribute Information
  +	rx_id 
      +	__Meaning__: The unique ID for the prescription
      +	__Variable Type__: character
      +	__Possible values__:  any number from 1 to infinity
  +	doctor_id 
      +	__Meaning__: The unique ID of a  doctor 
      +	__Variable Type__: character
      +	__Possible values__:  any number from 1 to infinity
  +	drug_name 
      +	__Meaning__: Name of the drug. Could be brand or generic name.
      +	__Variable Type__: character
      +	__Possible values__:  'OXYCODONE' , 'APIXABAN', 'FEELA-GOOD',etc
  +	is_pain_drug 
      +	__Meaning__: Is the drug used for pain or not. 1 if yes. 0 if no.
      +	__Variable Type__: integer
      +	__Possible values__:  1 or 0     

  +	is_target_drug 
      +	__Meaning__: Is the drug being prescribed the drug of interest? 1 if yes. 0 if no.
      +	__Variable Type__: integer
      +	__Possible values__:  1 or 0

  +	has_seen_msg 
      +	__Meaning__: Did the doctor see the campaign message?  1 if yes. 0 if no.
      +	__Variable Type__: integer
      +	__Possible values__:   1 or 0

# Table of Contents
1. Import Packages
2. Data load and EDA
3. Graphs
4. T-test assumptions
5. T-test
6. Conclusion

# 1.Import Packages
```{r ,warning=FALSE ,error=FALSE ,message=FALSE ,results='hide'}

library(tidyverse)
library(data.table)
library(openxlsx)
```

utility function
```{r}

factorize_str <- function (dt){
  #Purpose:Temporary converts a datatable's character datatypes into factor. This way when summarize function is used, it lists out the counts of the categorical variables
  #input:(datatable)
  #output:(datatable) returns a copy of the datatable but with all the character datatypes as factors
  
  #Making a copy to not affect the original datatable
  dt_copy <- copy(dt)

  for (col_name in colnames(dt_copy)){
    if (is.character(dt_copy[[col_name]])){
        set(dt_copy, j = col_name, value = as.factor(dt_copy[[col_name]]))
    }
  }
  
    
  dt_copy
}
```


# 2. Data load and EDA
```{r}
#Read in data
df <- openxlsx::read.xlsx('dataset.xlsx')

#Convert to datatable
dt <- as.data.table(df)

#changing datatypes
dt[ , doctor_id:=as.character(doctor_id)]
dt[ , rx_id:=as.character(rx_id)]

#Adding a column to help out with analysis
dt[ , rx_event :=1]

head(dt)
```


```{r}
#view summary of dataset
summary(factorize_str(dt))
glimpse(dt)

```

```{r}
#adding some columns to help out with graph generation
dt[ , intervention_group := ifelse(has_seen_msg ==1 ,'treatment','control')]
dt[ , categorical_target_value := ifelse(is_target_drug ==1 ,'target','not_target')]
dt[ , categorical_pain_value := ifelse(is_pain_drug ==1 ,'pain','not_pain')]
head(dt)

```

```{r}
#Check if there are any NAs
round(colSums(is.na(dt)) *100/nrow(dt),1)

```


```{r}
#check if a doctor can be in the treatment and the control group
dt[ , uniqueN(intervention_group) , by ='doctor_id'][V1>1]
```

```{r}
#Check if there are duplicate prescriptions
dt[ , .(count=.N) , by='rx_id'][count>1]

```


```{r}
#aggregate info at doctor level
doctor_agg_dt <- dt [ , .(group=max(intervention_group ), 
                          rx_cnt=.N , 
                          is_pain_drug_pct=round(sum(is_pain_drug)*100/.N,1),
                          is_target_drug_pct=round(sum(is_target_drug)*100/.N,1)),
                      by='doctor_id']
doctor_agg_dt
```


# 3.Graphs
```{r}
# Visualize the proportion of campaign drug being  prescribed between the treatment and control group   
ggplot(dt, aes (fill=categorical_target_value ,y=rx_event, x=intervention_group)) + 
    geom_bar(position="fill", stat="identity") + 
  labs( title = "Treatment vs Control:Proportions of Campaign Drug Prescriptions", y = "rx_events_proportion")
```
<br/>
- Insight: Seems like the treatment group prescribed more of the campaign drug. Maybe the promotion was effective???


```{r}
# See if the treatment and control group have similiar behavior in tendency of prescribing  pain related drugs
ggplot(dt, aes (fill=categorical_pain_value ,y=rx_event, x=intervention_group)) + 
    geom_bar(position="fill", stat="identity") +
  labs( title = "Treatment vs Control:Proportions of Pain Drug Prescriptions", y = "rx_events_proportion")
```
<br/>
- Insight: Seems like both groups have similar tendency to prescribe pain drugs. This is important attribute.This could be a confounding variable if it wasn't balance. We would need to do things to balance this covariate like propensity score weighting/matching.

```{r}
# Distribution of doctors' proportion of campaign drugs being prescribed
ggplot(doctor_agg_dt, aes(x = is_target_drug_pct , fill=group)) + 
  geom_density(alpha = 0.5)
```
<br/>
- Insight #1: Looks almost normally distributed. This is required for T-test.
- Insight #2: Looks like there could be a difference because the distributions aren't completely overlapping.

```{r}
# Distribution of doctors' proportion of pain prescriptions being prescribed
ggplot(doctor_agg_dt, aes(x = is_pain_drug_pct , fill=group)) + 
  geom_density(alpha = 0.5)
```
<br/>
- Insight:Lots of overlap. This matches the bar graphs from before, supporting the evidence that this covariate 'is_pain_drug_pct' is mostly balanced between two groups



```{r}

#See statistics at study group level
group_agg <- doctor_agg_dt[, .(rx_cnt = .N, 
                                is_pain_drug_pct = mean(is_pain_drug_pct),
                                is_target_drug_pct = mean(is_target_drug_pct)),
                            by = 'group']

group_agg
```
<br/>
- Insight: Seems like treatment group has much higher proportion of their prescriptions being related to the campaign




# 4.T-test assumptions

## Randomly Sample

It was fake data so probably was not randomly created. However, humans have good imagination skills so pretend this was randomly generated.

## independence

It was fake data so lets imagined each doctor didn't really talk to each other. So doctors can't influence each others behavior.


## Normalilty

```{r}

#  density graph
ggplot(doctor_agg_dt, aes(x = is_target_drug_pct , fill=group)) + 
  geom_density(alpha = 0.5)
#Insight:Not perfect but sort of close to normal distribution for both groups

```


## Similiar variances
```{r}
doctor_agg_dt[ , var(is_target_drug_pct) , by ='group']
#Insight:The ratio is less than 4 so this passes the variances test
```

# 5.T-test
```{r}

test_group_data <- doctor_agg_dt[group=='treatment']$is_target_drug_pct
control_group_data <- doctor_agg_dt[group=='control']$is_target_drug_pct

t.test(test_group_data, control_group_data, var.equal=TRUE)
```

# 6.conclusion

The campaign had a positive relative lift of 94%[(43.1-22.2)/(22.2)] for Feela-Good drug. This was statistically significant with a p-value < 0.001. The A/B testing showed that the campaign intervention was very effective in promoting the new drug. We should roll out more campaigns to doctors that share similar attributes to the test/control group.



