

#import libraries
```{r}

library(tidyverse)
library(data.table)
library(openxlsx)




```

#utility function
```{r}
#Temporary converts a datatable's character datatypes into factor. This way when summarize function is used, it lists out the counts of the categorical variables
factorize_str <- function (dt){
  #Making a copy to not affect the original datatable
  dt_copy <- copy(dt)

  for (col_name in colnames(dt_copy)){
    if (is.character(dt_copy[[col_name]])){
        set(dt_copy, j = col_name, value = as.factor(dt_copy[[col_name]]))
    }
  }
  
    
  dt_copy
}
```


#Load the Data

```{r}


df <- openxlsx::read.xlsx('dataset.xlsx')

dt <- as.data.table(df)

dt[ , doctor_id:=as.character(doctor_id)]
dt[ , rx_id:=as.character(rx_id)]
dt[ , rx_event :=1]

dt



```


```{r}


#view summary of dataset
summary(factorize_str(dt))
glimpse(dt)

```

```{r}

dt[ , intervention_group := ifelse(has_seen_msg ==1 ,'treatment','control')]
dt[ , categorical_target_value := ifelse(is_target_drug ==1 ,'target','not_target')]
dt[ , categorical_pain_value := ifelse(is_pain_drug ==1 ,'pain','not_pain')]
dt

```

```{r}
#Check if there are any NAs
round(colSums(is.na(dt)) *100/nrow(dt),1)



```


```{r}

#check if a doctor can be in the treatment and the control group

dt[ , uniqueN(intervention_group) , by ='doctor_id'][V1>1]





```

```{r}
#Check if there are duplicate prescriptions

dt[ , .(count=.N) , by='rx_id'][count>1]

```


```{r}

doctor_agg_dt <- dt [ , .(group=max(intervention_group ), 
                          rx_cnt=.N , 
                          is_pain_drug_pct=round(sum(is_pain_drug)*100/.N,1),
                          is_target_drug_pct=round(sum(is_target_drug)*100/.N,1)),
                      by='doctor_id']
doctor_agg_dt
```


```{r}

ggplot(doctor_agg_dt, aes(x=doctor_id, y=is_target_drug_pct,color=group)) +
  geom_point() 


```

```{r}




ggplot(dt, aes (fill=categorical_target_value ,y=rx_event, x=intervention_group)) + 
    geom_bar(position="fill", stat="identity")


```
```{r}
ggplot(dt, aes (fill=categorical_pain_value ,y=rx_event, x=intervention_group)) + 
    geom_bar(position="fill", stat="identity")

```





```{r}


# Histogram with kernel density
ggplot(doctor_agg_dt, aes(x = is_target_drug_pct)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density()
```

```{r}
# Histogram with kernel density
ggplot(doctor_agg_dt, aes(x = is_target_drug_pct , fill=group)) + 
  geom_histogram(aes(y = ..density..)) +
  geom_density()


```

```{r}

# Histogram with kernel density
ggplot(doctor_agg_dt, aes(x = is_target_drug_pct , fill=group)) + 
  
  geom_density(alpha = 0.5)


```

```{r}
# Histogram with kernel density
ggplot(doctor_agg_dt, aes(x = is_pain_drug_pct , fill=group)) + 
  
  geom_density(alpha = 0.5)
```


```{r}


group_agg <- doctor_agg_dt[, .(rx_cnt = .N, 
                                is_pain_drug_pct = mean(is_pain_drug_pct),
                                is_target_drug_pct = mean(is_target_drug_pct)),
                            by = 'group']

group_agg
```


```{r}


group_agg2 <-   dt[, .(rx_cnt = .N, 
                                is_pain_drug_pct = round(mean(is_pain_drug)*100,1),
                                is_target_drug_pct = round(mean(is_target_drug)*100,1)),
                            by = 'intervention_group']

group_agg2
```



#T-test assumptions

##Randomly Sample

It was fake data so probably was not randomly created. However, humans have good imagination skills so pretend this was randomly generated.

## independence

It was fake data so lets imagined each doctor didn't really talk to each other. So doctors can't influence each others behavior.


##Normalilty

```{r}


# Histogram with kernel density
ggplot(doctor_agg_dt, aes(x = is_target_drug_pct , fill=group)) + 
  geom_density(alpha = 0.5)



#Not perfect but sort of close to normal distribution for both groups

```


##Similiar variances
```{r}

doctor_agg_dt[ , var(is_target_drug_pct) , by ='group']

#The ratio is less than 4 so this passes the variances test
```


```{r}
doctor_agg_dt
```


#T-test
```{r}

test_group_data <- doctor_agg_dt[group=='treatment']$is_target_drug_pct
control_group_data <- doctor_agg_dt[group=='control']$is_target_drug_pct

t.test(test_group_data, control_group_data, var.equal=TRUE)
```

#conclusion



